Relationships in the extremes and their influence on competition and coexistence

### 2022-03-01 paper/Paper.Rnw
edits to the Intro to reduce length

%Currently 148 of an allowed 150 words for Ecology Letters
\begin{abstract}
\noindent Species coexistence attracts wide 
interest in ecology. Modern coexistence
theory (MCT) identifies mechanisms of coexistence,
one of which, storage effects, hinges on relationships between fluctuations
in environmental and competitive pressures. However, such relationships
are typically measured using covariance, which does not account for 
the possibility that environmental and competitive influences may be more related to each other
when they are strong than when weak, or vice versa. 
Recent work showed that such `asymmetric tail associations' (ATAs) are common
between ecological variables, and are important for 
extinction risk, ecosystem stability, and other phenomena. 
We extend the MCT to account for ATAs. Storage
effects are decomposed to show the influence of ATAs. Simulations of a simple model
and an empirical example using diatoms
illustrate that ATA influences can be comparable in magnitude to other
mechanisms of coexistence, and that ATAs can make the 
difference between species coexistence and competitive exclusion. 
\end{abstract}

@@

%Intro to competition, coexistence, MCT, etc.
Competition and the maintenance of biodiversity are research topics at the core of modern ecology 
\citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}.
The competitive exclusion principle states 
that two species competing for the same limiting resource cannot coexist at constant densities - one will instead outcompete 
the other and drive it to extinction \citep{Hardin_1960}.
Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
framework for understanding how species nevertheless coexist, and biodiversity is maintained, in real ecosystems. 
Coexistence occurs through stabilizing factors and niche 
differentiation, in spite of the conclusions drawn for idealized scenarios by the competitive exclusion principle. 
MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``relative nonlinearity
in competition'' and ``storage effects'' \citep{Chesson_1994} which will be familiar to many readers. MCT has been
applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2010, Adler_2006}, 
tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Descamps_2005, Caceres_1997}, and corals \citep{Alvarez-Noriega_2020}. 
Although MCT is mathematically complex to many ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.

%Intro to ATAs
Every ecological discipline involves the study of relationships among biological and environmental variables,
but standard methods for examining the association between two variables provide only a small 
portion of the overall information content of the
relationship \citep{Ghosh_2020_AER,Ghosh_2020_Eco,Ghosh_2020_ecolevol,Ghosh_2021}. For instance, 
the three pairs of variables in Fig. \ref{fig:noise} all have the same
Pearson correlation, but nevertheless demonstrate very different types of association. Other metrics 
of association are also often used, in addition to correlations
(e.g., various forms of regression), but correlations are extremely common, and other
methods also reveal only a small portion of the information content of the relationship between variables \citep{Nelsen2006}.
The variables in Fig. \ref{fig:noise}a,c 
exhibit what we have previously referred to as ``asymmetric tail association'' [ATA; \cite{Ghosh_2020_AER}], i.e., 
they are more strongly related to each
other in the left (Fig. \ref{fig:noise}a, termed ``left-tail association'') or right (Fig. \ref{fig:noise}c,
termed ``right-tail association'') portions of the variables' distributions. Although ATAs 
and related ideas had previously been seldom 
considered in ecology (but see \citealp{Anderson_2019, deValpine_2014, Popovic_2019}), 
our recent work has shown that, in fact, ATAs are extremely common in ecology \citep{Ghosh_2020_AER}, and can influence 
many ecological phenomena, including extinction risk \citep{Ghosh_2020_Eco}, Taylor's law \citep{Ghosh_2020_AER}, 
synchrony, and the temporal stability of ecosystem functioning \citep{Ghosh_2020_ecolevol,Ghosh_2021}.

%Intro to storage effects and why they might be influenced by ATAs
Although ATAs are now known to be common and important, the implications of ATAs for competition and species coexistence 
have, to our knowledge, not previously been considered; doing so is the overarching mission of this study. 
It is straightforward to see, intuitively, 
why ATAs may influence one mechanism of coexistence in particular: storage effects. We review, conceptually, 
what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and we explain why ATAs may be relevant to them.
Storage effects allow for species to fluctuate over time while escaping extinction because storage effects provide a relative benefit to
whichever species is currently rare; this is the characteristic that makes storage effects a so-called \emph{stabilizing} mechanism 
of coexistence \citep{Chesson_2000}.
Storage effects can occur if three conditions are met \citep{Chesson_1994}. First, there must be positive 
covariation between the fluctuating quality of the environment (E)
for a species and the strength of competition (C) experienced by that species (called EC covariance). 
Second, growth responses to the environment must be species-specific, so that distinct species do not find the 
same environmental conditions to be optimal. Third, a buffering mechanism must exist whereby species limit the influence of competition
in times when the environment is poor. Dormant stages such as seed banks are an
example buffering mechanism \citep{Chesson_2012}. 
When present, these three factors combine so that EC covariance provides a relative benefit to whatever species are currently rare.
Importantly, storage effects can be quantified by measuring to what extent EC covariance contributes to differences between 
invader and resident growth rates \citep{Chesson_1994,Ellner_2016}.
Thus ATAs may, in principle, influence storage effects simply because storage effects hinge on EC covariation, and ATAs, when they 
occur between environment and competition variables, can be an aspect of EC covariation.

@@

%Climate change may alter ATAs of coexistence-relevant ATAs
An additional motivation for quantifying the importance of ATAs for species coexistence is that there is reason to believe climate change may
alter ATAs of environmental variables which impact coexistence, as follows.
It is well known that climate change can alter the intensity, duration, extent and frequency of extreme environmental events, including
heat waves, floods, hurricanes and fires \citep{Jentsch_2007, Meehl_2004, Ummenhofer_2017, Lyon2019, Keelings2020}.  
If extreme events are also longer lasting, then the nature of the covariation between phenologically separated environmental signals 
may be modified by changes in extreme climatic events; in particular, ATAs between such variables may be altered. 
For instance, imagine a case in which early spring precipitation influences the growth of a plant species, A,
whereas late spring precipitation influences plant species B. If these species are competitors, then EC covariation
contributing to their coexistence, or failure to coexist, can boil down to covariation 
between early and late spring precipitation. Because longer lasting weather events are more likely to contribute
to both early and late spring rainfall variables, a tendency for spring rainfall events to become both more extreme and 
longer lasting increases the degree of right-tail association between early- and late-spring rainfall. 
Likewise, a tendency for spring droughts to become both more intense and longer lasting can increase
left-tail association between these variables. As we will show in this study, changes in ATAs can then
result in different competitive outcomes between species. Although we focus here on temporal coexistence
mechanisms, ATAs between environmental variables measured in different places may also be altered 
by climate change, if more intensive climatic events are also more spatially extensive.

%make the above abstract ideas more concrete
To make the above ideas more concrete, we briefly mention examples from prior work to 
which the ideas may apply. 
\cite{Chesson_2012} provides two examples of competitive plant communities in which different species
are likely influenced differently by phenologically distict environmental variables: an Australian 
grass-forb community, and an Arizona community of desert annuals. 
The communities follow the criteria required for temporal storage effects to operate \citep{Chesson_2012}. In both 
examples, distinct species have different periods of germination, competition is influenced by resource 
consumption, and population growth is buffered by overlapping generations or via seed banks.
Temporal differences between species in germination rates suggest the possibility that 
population growth of distinct species may be driven by rainfall signals in distinct parts
of the growing season. Thus all the ingredients seem to be present by which climatic changes may
alter ATAs between environmental variables which are relevant for the coexistence
of the example species, motivating our efforts, here, to perform a detailed study of how ATAs can influence competition.

%Review of lit on changes in means and variances of climate signals and associated impacts on coexistence
Although, to our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
not been studied, the importance for coexistence of changes 
in means and variability of environmental variables has been studied in many systems. We briefly 
review some of that literature.
\cite{diPaola_2012} concluded that coexistence of drought sensitive,
flood resistant species with flood sensitive, drought resistant species in 
Mediteranean coastal swamp forests of central Italy would be jeopardized
by projected changes in mean rainfall. 
\cite{Descombes_2020} found that if warming trends promote faster expansion toward 
higher altitudes of herbivore populations than of plant populations, newly arrived
herbivores at high altitudes can inflate plant species richness by reducing dominant plant species 
in favor of subordinates. Contrastingly, \cite{Usinowicz_2021} concluded, also for
alpine plants, that upward range expansion under warming trends would be good for the growth and survival of 
all species they considered, but that changes in competition would more than offset these direct benefits. 
Numerous studies take an ``event-focused'' 
perspective \citep{Jentsch_2007}, examining how increased climatic variability, in the
form of extreme weather events, can alter competitive interactions. For instance, 
\cite{White_2001} showed that single extreme heat events can alter competition 
between C3 and C4 plants. And \cite{Parepa_2013} showed that extreme variability in soil
nutrient availability for plants, mediated by soil moisture content, can alter 
the potential for invasive species to enter an ecosystem and establish, a process that involves
the invader overcoming competition from resident species. 
Using MCT, it has been demonstrated that increased variability can promote coexistence in plant 
communities, via storage effects \citep{Adler_2006,  Angert_2009}.
Changes in variability and in extreme events are distinct from the changes in ATAs we 
consider here, because ATAs pertain to \emph{relationships} between environmental and other
variables in the extremes, which are distinct from and can be altered independently
of changes in the extremes of the univariate marginal distributions themselves \citep{Nelsen2006,Ghosh_2020_AER}.

@@

%Goal para
We previously demonstrated \citep{Ghosh_2020_AER} that ATAs are common in nature, and we argued above 
that it is plausible both that climate change may alter environmental ATAs, and that ATAs may influence species 
coexistence. Thus we seek to explore the consequences of ATAs for competition, coexistence, and the maintenance of 
biodiversity, by means of the following specific goals. 
G1) To augment MCT to show formally how ATA plays a role in coexistence. 
MCT decomposes a measure of coexistence into components due to each of several mechanisms. 
We address how ATA contributes to these components. 
We here examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; 
spatial variation can be considered in future work.
G2) To explore, using the simple lottery model, the relative magnitudes of the contributions to coexistence that come
from ATAs compared to other, previously known mechanisms of coexistence. This is to help 
illuminate whether ATA contributions to coexistence 
are likely to be negligible or meaningfully large. 
G3) To explore, using the lottery model, whether species can coexist with
ATAs but not without, or vice versa. This is to begin to illuminate whether ATA contributions to biodiversity maintenance 
may be negligible or meaningfully large.
G4) To provide an empirical example of species coexistence which highlights the role ATAs can play.
In the Discussion, we revisit the possibility that climate change will alter
ATAs of environmental variables and thereby influence competition and species coexistence.
Overall, our study presents a new mechanism of species coexistence which may be influenced by climate change, 
and a means of understanding its theoretical importance and its importance for real ecosystems.

@@

@@

For a general model, it is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
of species $i$ can be written as an increasing function of
an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
competitive pressure on species $i$, $C_i(t)$. As for the lottery model example (Introduction), we here assume, for simplicity, 
that the $E_i(t)$ are iid through time, and likewise for the
$C_i(t)$. Though this assumption can be circumvented, we make the assumption because it simplifies the
presentation of new ideas.

@@

\begin{figure}
\includegraphics[width=.75\textwidth]{../results_figs/fig2.pdf}
\caption{Our simulated lottery-model populations influenced by asymmetric tail associations (ATAs) recovered from low densities less often and 
spent more time in a state of one-species dominance, demonstrating that ATAs can influence aspects of 
species coexistence. Panel (a) shows species 1, from the lottery 
model subject to environmental noise with symmetric tail associations (Fig. \ref{fig:noise}b). Panel (b) shows species 1 from the 
same model with the same parameters, but subject to noise with left-tail association (Fig. \ref{fig:noise}a).
Both simulations were run for \Sexpr{M} time steps, though only the first $1000$ are shown. Periods of 
``noticeable coexistence'' of the species, defined to be periods when both species each comprise at least $5\%$ of 
total abundance, are marked in grey. The average duration of periods of noticeable coexistence 
was \Sexpr{fig2$coSym} time steps for panel (a) and \Sexpr{fig2$coATA} for panel (b), over the full length of the simulations; 
whereas the average duration of periods of
one-species dominance was \Sexpr{fig2$domSym} time steps for panel (a) and \Sexpr{fig2$domATA} for (b). Thus, while periods of 
noticeable coexistence were of similar duration for the two scenarios, periods of one-species dominance were much longer,
on average, for the ATA case. Parameters were $\mu_1 = \Sexpr{fig2$params[1] + fig2$params[1]}$ and $\mu_2 = \Sexpr{-(fig2$params[1])}$, $\sigma = \Sexpr{fig2$params[2]}$, and $\delta = \Sexpr{fig2$params[3]}$. See Introduction and Fig. \ref{fig:noise} caption for model details. 
\label{fig:pedagog}}
\end{figure}

>>>>

%Currently 150 of an allowed 150 words for Ecology Letters
\begin{abstract}
\noindent Species coexistence attracts wide 
interest in ecology. Modern coexistence
theory (MCT) identifies coexistence mechanisms,
one of which, storage effects, hinges on relationships between fluctuations
in environmental and competitive pressures. However, such relationships
are typically measured using covariance, which does not account for 
the possibility that environment and competition may be more related to each other
when they are strong than when weak, or vice versa. 
Recent work showed that such `asymmetric tail associations' (ATAs) are common
between ecological variables, and are important for 
extinction risk, ecosystem stability, and other phenomena. 
We extend the MCT, decomposing storage
effects to show the influence of ATAs. Analysis of a simple model
and an empirical example using diatoms
illustrate that ATA influences can be comparable in magnitude to other
mechanisms of coexistence, and that ATAs can make the 
difference between species coexistence and competitive exclusion.
ATA influences are an important new mechanism of coexistence.
\end{abstract}

@@

%Intro to competition, coexistence, MCT, etc.
Competition and the maintenance of biodiversity are research topics at the core of modern ecology 
\citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}.
The competitive exclusion principle states 
that two species competing for the same limiting resource cannot coexist at constant densities - one will instead outcompete 
the other and drive it extinct \citep{Hardin_1960}.
Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
framework for understanding how species nevertheless coexist, and biodiversity is maintained, in real ecosystems. 
MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``relative nonlinearity
in competition'' and ``storage effects'' \citep{Chesson_1994} which will be familiar to many readers. MCT has been
applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
tropical trees \citep{Usinowicz_2012}, plankton \citep{Descamps_2005, Caceres_1997, Narawani_2013}, and corals \citep{Alvarez-Noriega_2020}. 
Although MCT is mathematically complex to many ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.

%Intro to ATAs
Every ecological discipline involves the study of relationships among biological and environmental variables,
but standard methods for examining the association between two variables provide only a small 
portion of the overall information content of the
relationship \citep{Ghosh_2020_AER,Ghosh_2020_Eco,Ghosh_2020_ecolevol,Ghosh_2021}. For instance, 
the three pairs of variables in Fig. \ref{fig:noise} all have the same
Pearson correlation, but nevertheless demonstrate very different types of association. Other metrics 
of association are also often used, in addition to correlations
(e.g., various forms of regression), but correlations are extremely common, and other
methods also reveal only a small portion of the information content of the relationship between variables \citep{Nelsen2006}.
The variables in Fig. \ref{fig:noise}a,c 
exhibit what we have previously referred to as ``asymmetric tail association'' [ATA; \cite{Ghosh_2020_AER}], i.e., 
they are more strongly related to each
other in the left (Fig. \ref{fig:noise}a, termed ``left-tail association'') or right (Fig. \ref{fig:noise}c,
termed ``right-tail association'') portions of the variables' distributions. Although ATAs 
and related ideas had previously been seldom 
considered in ecology (but see \citealp{deValpine_2014, Anderson_2019, Popovic_2019}), 
our recent work shows that ATAs are extremely common \citep{Ghosh_2020_AER}, and can influence 
many phenomena, including extinction risk \citep{Ghosh_2020_Eco}, Taylor's law \citep{Ghosh_2020_AER}, 
and stability of ecosystem functioning \citep{Ghosh_2020_ecolevol,Ghosh_2021}.
Although ATAs are now known to be important, the implications of ATAs for competition and species coexistence 
have, to our knowledge, not previously been considered; doing so is the overarching mission of this study.

%Intro to storage effects and why they might be influenced by ATAs
It is straightforward to see, intuitively, 
why ATAs may influence one mechanism of coexistence in particular: storage effects. We review, conceptually, 
what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs may be relevant to them.
Storage effects allow for species to fluctuate over time while escaping extinction because storage effects provide a relative benefit to
whichever species is currently rare; this is the characteristic that makes storage effects a so-called \emph{stabilizing} mechanism 
of coexistence \citep{Chesson_2000}.
Storage effects can occur if three conditions are met \citep{Chesson_1994}. First, there must be positive 
covariation between the fluctuating quality of the environment (E)
for a species and the strength of competition (C) experienced by that species (called EC covariance). 
Second, growth responses to the environment must be species-specific, so that distinct species do not find the 
same environmental conditions to be optimal. Third, a buffering mechanism must exist whereby species limit the influence of competition
in times when the environment is poor. Dormant stages such as seed banks are an
example buffering mechanism \citep{Chesson_2012}. 
When present, these three factors combine so that EC covariance provides a relative benefit to whatever species is currently rare.
Importantly, storage effects can be quantified by measuring to what extent EC covariance contributes to differences between 
invader and resident growth rates \citep{Chesson_1994,Ellner_2016}.
Thus ATAs may, in principle, influence storage effects simply because storage effects hinge on EC covariation, and ATAs, when they 
occur between environment and competition variables, can be an aspect of EC covariation.

%***DAN: Go though and make sure all citation callouts list papers in date order, olderst to newest,
%except when multiple pubs by the same first author are listed they are all together.

@@

%Climate change may alter ATAs of coexistence-relevant ATAs
An additional motivation for quantifying the importance of ATAs for species coexistence is that there is reason to believe climate change may
alter ATAs of environmental variables which impact coexistence, as follows.
It is well known that climate change can alter the intensity, duration, extent and frequency of extreme environmental events, including
heat waves, floods, hurricanes and fires \citep{Meehl_2004, Jentsch_2007, Ummenhofer_2017, Lyon2019, Keelings2020}.  
If extreme events are also longer lasting, then the nature of the covariation between phenologically separated environmental signals 
may be modified by changes in extreme climatic events; in particular, ATAs between such variables may be altered. 
For instance, imagine a case in which early spring precipitation influences the growth of a plant species, A,
whereas late spring precipitation influences plant species B. If these species are competitors, then EC covariation
contributing to their coexistence, or failure to coexist, can boil down to covariation 
between early and late spring precipitation. Because longer lasting weather events are more likely to contribute
to both early and late spring rainfall variables, a tendency for spring rainfall events to become both more extreme and 
longer lasting increases the degree of right-tail association between early- and late-spring rainfall. 
Likewise, a tendency for spring droughts to become both more intense and longer lasting can increase
left-tail association between these variables. 
As we will show in this study, changes in ATAs can then
result in different competitive outcomes between species.
\cite{Chesson_2012} provides examples from an Australian grass-forb community and from an Arizona 
community of desert annuals that seem to have all the ingredients necessary to make effects of the
type we describe here a possibility.
To our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
not been studied, but the importance for coexistence of changes 
in means and variability of environmental variables has been studied in many systems 
[e.g., \cite{White_2001, Adler_2006, Jentsch_2007, Angert_2009, diPaola_2012, Parepa_2013, Descombes_2020, Usinowicz_2021}]. 
Changes in variability and in extreme events are distinct from the changes in ATAs we 
consider here, because ATAs pertain to \emph{relationships} between environmental and other
variables in the extremes, which are distinct from and can be altered independently
of changes in the extremes of the univariate marginal distributions themselves \citep{Nelsen2006,Ghosh_2020_AER}.

%make the above abstract ideas more concrete - 
%***DAN: This para commented out for length, added a sentence about the systems to the above paragraph to take the place of this text
%To make the above ideas more concrete, we briefly mention examples from prior work to 
%which the ideas may apply. 
%\cite{Chesson_2012} provides two examples of competitive plant communities in which different species
%are likely influenced differently by phenologically distict environmental variables: an Australian 
%grass-forb community, and an Arizona community of desert annuals. 
%The communities follow the criteria required for temporal storage effects to operate \citep{Chesson_2012}. In both 
%examples, distinct species have different periods of germination, competition is influenced by resource 
%consumption, and population growth is buffered by overlapping generations or via seed banks.
%Temporal differences between species in germination rates suggest the possibility that 
%population growth of distinct species may be driven by rainfall signals in distinct parts
%of the growing season. Thus all the ingredients seem to be present by which climatic changes may
%alter ATAs between environmental variables which are relevant for the coexistence
%of the example species, motivating our efforts, here, to perform a detailed study of how ATAs can influence competition.

%Review of lit on changes in means and variances of climate signals and associated impacts on coexistence  
%***DAN: this para commented out for length, added a couple of the sentences to an above para to substitute
%Although, to our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
%not been studied, the importance for coexistence of changes 
%in means and variability of environmental variables has been studied in many systems. We briefly 
%review some of that literature.
%\cite{diPaola_2012} concluded that coexistence of drought sensitive,
%flood resistant species with flood sensitive, drought resistant species in 
%Mediteranean coastal swamp forests of central Italy would be jeopardized
%by projected changes in mean rainfall. 
%\cite{Descombes_2020} found that if warming trends promote faster expansion toward 
%higher altitudes of herbivore populations than of plant populations, newly arrived
%herbivores at high altitudes can inflate plant species richness by reducing dominant plant species 
%in favor of subordinates. Contrastingly, \cite{Usinowicz_2021} concluded, also for
%alpine plants, that upward range expansion under warming trends would be good for the growth and survival of 
%all species they considered, but that changes in competition would more than offset these direct benefits. 
%Numerous studies take an ``event-focused'' 
%perspective \citep{Jentsch_2007}, examining how increased climatic variability, in the
%form of extreme weather events, can alter competitive interactions. For instance, 
%\cite{White_2001} showed that single extreme heat events can alter competition 
%between C3 and C4 plants. And \cite{Parepa_2013} showed that extreme variability in soil
%nutrient availability for plants, mediated by soil moisture content, can alter 
%the potential for invasive species to enter an ecosystem and establish, a process that involves
%the invader overcoming competition from resident species. 
%Using MCT, it has been demonstrated that increased variability can promote coexistence in plant 
%communities, via storage effects \citep{Adler_2006,  Angert_2009}.
%Changes in variability and in extreme events are distinct from the changes in ATAs we 
%consider here, because ATAs pertain to \emph{relationships} between environmental and other
%variables in the extremes, which are distinct from and can be altered independently
%of changes in the extremes of the univariate marginal distributions themselves \citep{Nelsen2006,Ghosh_2020_AER}.

@@

%Goal para
We previously demonstrated \citep{Ghosh_2020_AER} that ATAs are common in nature, and we argued above 
that it is plausible both that climate change may alter environmental ATAs, and that ATAs may influence species 
coexistence. Thus we seek to explore the consequences of ATAs for competition, coexistence, and the maintenance of 
biodiversity, by means of the following specific goals. 
G1) To augment MCT to show formally how ATAs play a role in coexistence. 
MCT decomposes a measure of coexistence into components due to each of several mechanisms. 
We address how ATAs contribute to these components. 
We here examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; 
spatial variation can be considered in future work.
G2) To explore, using the simple lottery model, the relative magnitudes of the contributions to coexistence that come
from ATAs compared to other, previously known mechanisms of coexistence. This is to help 
illuminate whether ATA contributions to coexistence 
are likely to be negligible or meaningfully large. 
G3) To explore, using the lottery model, whether species can coexist with
ATAs but not without, or vice versa. This is to begin to illuminate whether ATA contributions to biodiversity maintenance 
may be negligible or meaningfully large.
G4) To provide an empirical example of species coexistence which highlights the role ATAs can play.
In the Discussion, we revisit the possibility that climate change will alter
ATAs of environmental variables and thereby influence competition and species coexistence.
%***DAN: Make sure we end up having such material in the Discussion
Overall, our study presents a new mechanism of species coexistence which may be influenced by climate change, 
and a means of understanding its theoretical and empirical importance.

@@

%***DAN: consider sending the lottery model components of this section to sup mat, for length reasons.

@@

For a general model, it is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
of species $i$ can be written as an increasing function of
an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
competitive pressure on species $i$, $C_i(t)$. We assume, for simplicity, 
that the $E_i(t)$ are iid through time, and likewise for the
$C_i(t)$. Though this assumption can be circumvented, we make the assumption because it simplifies the
presentation of new ideas.

%***DAN: I think we should replace the terminology "invasion growth rate" and "IGR" by "growth when rare" and "GWR",
%throughout the paper. Has to be done in the main text and sup mat and the figures themselves (IGR is often labeled on 
%the figure panels themselves).

@@

\begin{figure}
\includegraphics[width=.75\textwidth]{../results_figs/fig2.pdf}
\caption{Our simulated lottery-model populations influenced by asymmetric tail associations (ATAs) recovered 
from low densities less often and 
spent more time in a state of one-species dominance, demonstrating that ATAs can influence aspects of 
species coexistence. Panel (a) shows species 1 from the lottery 
model subject to environmental noise with symmetric tail associations (Fig. \ref{fig:noise}b). Panel (b) shows species 1 from the 
same model with the same parameters, but subject to noise with left-tail association (Fig. \ref{fig:noise}a).
Both simulations were run for \Sexpr{M} time steps, though only the first $1000$ are shown. Periods of 
``noticeable coexistence'' of the species, defined to be periods when each species comprises at least $5\%$ of 
total abundance, are marked in grey. The average duration of periods of noticeable coexistence 
was \Sexpr{fig2$coSym} time steps for panel (a) and \Sexpr{fig2$coATA} for panel (b), over the full length of the simulations; 
whereas the average duration of periods of
one-species dominance was \Sexpr{fig2$domSym} time steps for panel (a) and \Sexpr{fig2$domATA} for (b). Thus, while periods of 
noticeable coexistence were of similar duration for the two scenarios, periods of one-species dominance were much longer,
on average, for the ATA case. Parameters were $\mu_1 = \Sexpr{fig2$params[1] + fig2$params[1]}$, 
$\mu_2 = \Sexpr{-(fig2$params[1])}$, $\sigma = \Sexpr{fig2$params[2]}$, and $\delta = \Sexpr{fig2$params[3]}$. 
See Introduction and Fig. \ref{fig:noise} caption for model details. 
\label{fig:pedagog}}
\end{figure}
